---
title: 'Simulation: All Plots'
author: "Yiqi Liu & Francesca Molinari"
date: '`r Sys.Date()`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, echo = TRUE)
library(gridExtra)
library(grid)
library(ggplot2)
library(latex2exp)
library(concaveman)
source("../../all-func.R")

set.seed(0)
```

## Data generating process

$$
X \equiv (X_1, X_2, ..., X_{20})\in\mathbb{R}^{20}\\
\text{for } j \in \{1,4,...,20\},~X_j \stackrel{d}{\sim} Normal(0,1) \text{ truncated to } [-3,3],\\
X_2 \stackrel{d}{\sim} Unif(0,1),~X_3\stackrel{d}{\sim}Beta(2,2),~G \stackrel{d}{\sim} Bern\left(p=0.6\right)
$$

For a group-balanced DGP, we generate $$
Y \,|\, G, X\,\,\stackrel{d}{\sim} Bern\left(p=\frac{G}{1+e^{-(X_1+X_2+0.5X_3)}}+\frac{(1-G)}{1+e^{-(-X_1-0.5X_2+X_4)}}\right)
$$

For an $r$-skewed DGP, we generate $$
Y \,|\, G, X\,\,\stackrel{d}{\sim} Bern\left(p=\frac{G}{1+e^{-2(X_1+X_2+X_3)}}+\frac{(1-G)}{1+e^{-0.7(X_1+0.5X_2+0.6X_4)}}\right)
$$ We consider testing, for both of the group-balanced and r-skewed DGPs, whether there is an LDA to the following algorithm, `alg_logit`, which is a logistic regression trained on a sample consisted of data from both DGPs.

```{r}
train_data1 <- testDGP(5000)
train_data2 <- testDGP(5000)

train_X <- rbind(train_data1$X, train_data2$X)
train_Y <- append(train_data1$Y_balance, train_data2$Y_rskew)

train_data <- data.frame(cbind(train_X, train_Y))
colnames(train_data) <- append(sapply(1:20, function(num){paste0("X",num)}), "Y")

alg_logit <- glm(Y ~ ., data = train_data, family=binomial())
```

We then take `alg_logit` as the given $a^*$, and simulate the true risk allocation $e^*$ induced by $a^*$ for each of the group-balanced and r-skewed DGP (plotted below as a green star).

```{r include=FALSE}
test_data <- testDGP(1e7)
eval_Y_balance <- test_data$Y_balance
eval_Y_rskew <- test_data$Y_rskew
eval_G <- test_data$G
eval_data_X <- data.frame(test_data$X)

colnames(eval_data_X) <- sapply(1:20, function(num){paste0("X",num)})

alg_pred <- predict.glm(alg_logit, eval_data_X, type="response")

true_mu_1 <- 0.6

logit_e_score_balance <- alg_pred*as.numeric(1 != eval_Y_balance)+(1-alg_pred)*as.numeric(0 != eval_Y_balance)
logit_e_balance <- matrix(c(mean(as.numeric(eval_G==1)*logit_e_score_balance/true_mu_1),
                            mean(as.numeric(eval_G==0)*logit_e_score_balance/(1-true_mu_1))), 
                          ncol=2)

logit_e_score_rskew <- alg_pred*as.numeric(1 != eval_Y_rskew)+(1-alg_pred)*as.numeric(0 != eval_Y_rskew)
logit_e_rskew <- matrix(c(mean(as.numeric(eval_G==1)*logit_e_score_rskew/true_mu_1),
                            mean(as.numeric(eval_G==0)*logit_e_score_rskew/(1-true_mu_1))), 
                          ncol=2)

num_q <- 500 # number of directions to check
grid_q <- t(as.matrix(sapply(seq(0, 2*pi, length.out=num_q), function(rad){c(cos(rad), sin(rad))})))

# grid of (q, v) used to simulate the feasible set
grid_q <- rbind(grid_q, grid_q)
grid_v <- data.matrix(cbind(c(rep(1, each=num_q), rep(0, each=num_q)), c(rep(0, each=num_q), rep(1, each=num_q))))

# directions (q, v) that yield coordinates of R and B
RB_q <- matrix(c(-1, 0, -1, 0, 0, -1, 0, -1), byrow=TRUE, ncol=2)
RB_v <- matrix(c(1, 0, 0, 1, 1, 0, 0, 1), byrow=TRUE, ncol=2)

q <- rbind(grid_q, RB_q)
v <- rbind(grid_v, RB_v)

test_X1 <- test_data$X[,1]
test_X2 <- test_data$X[,2]
test_X3 <- test_data$X[,3]
test_X4 <- test_data$X[,4]

# true theta_dg(X) for each X in test_data
true_mu_1 <- 0.6
true_theta_11_balance <- 1-1/(1+exp(-(test_X1+test_X2+0.5*test_X3)))
true_theta_01_balance <- 1/(1+exp(-(test_X1+test_X2+0.5*test_X3)))
true_theta_10_balance <- 1-1/(1+exp(-(-test_X1-0.5*test_X2+test_X4)))
true_theta_00_balance <- 1/(1+exp(-(-test_X1-0.5*test_X2+test_X4)))

true_theta_11_rskew <- 1-1/(1+exp(-2*(test_X1+test_X2+test_X3)))
true_theta_01_rskew <- 1/(1+exp(-2*(test_X1+test_X2+test_X3)))
true_theta_10_rskew <- 1-1/(1+exp(-0.7*(test_X1+0.5*test_X2+0.6*test_X4)))
true_theta_00_rskew <- 1/(1+exp(-0.7*(test_X1+0.5*test_X2+0.6*test_X4)))

## group-balanced config
true_sf_balance <- matrix(data=NA, ncol=1, nrow=nrow(q)) # initialize final output vector

diff_theta_b <- true_theta_10_balance-true_theta_00_balance # theta_1^b - theta_0^b
diff_theta_r <- true_theta_11_balance-true_theta_01_balance # theta_1^r - theta_0^r

for (i in 1:nrow(q)){
  # direction (q, v) in the current iteration
  q_cur <- q[i,]
  v_cur <- v[i,]
  inner_prod_q <- q_cur[1]*diff_theta_r + q_cur[2]*diff_theta_b # q'k(\theta)
  inner_prod_v <- v_cur[1]*diff_theta_r + v_cur[2]*diff_theta_b # v'k(\theta)
  sign_inner_prod <- as.numeric(inner_prod_q > 0) # 1(q'k(\theta) > 0)
  
  score <- v_cur[1]*true_theta_01_balance+v_cur[2]*true_theta_00_balance+
    inner_prod_v*as.numeric(sign_inner_prod>0)
  
  # then the simulated support function in the current direction (q_cur, v_cur) is the sample average of score over X
  true_sf_balance[i,] <- mean(score)
}

## SIMULATING TRUE F
f <- function(c){ # objective function
  mean((-1-c)*true_theta_01_balance+c*true_theta_00_balance+
         ((-1-c)*diff_theta_r + c*diff_theta_b)*
         as.numeric((-1-c)*diff_theta_r + c*diff_theta_b>0))
}

c <- torch_tensor(0, requires_grad=TRUE)
optim <- optim_adam(params=list(c), lr=0.1)
for (i in 1:10000) {
  optim$zero_grad()            # initialize
  loss <- f(c)                 # get loss at current c
  loss$backward()              # backpropagation
  optim$step()                 # descent
}

true_F_coord_balance <- -loss$item() # 0.415

## r-skewed config
true_sf_rskew <- matrix(data=NA, ncol=1, nrow=nrow(q)) # initialize final output vector

diff_theta_b <- true_theta_10_rskew-true_theta_00_rskew # theta_1^b - theta_0^b
diff_theta_r <- true_theta_11_rskew-true_theta_01_rskew # theta_1^r - theta_0^r

for (i in 1:nrow(q)){
  # direction (q, v) in the current iteration
  q_cur <- q[i,]
  v_cur <- v[i,]
  inner_prod_q <- q_cur[1]*diff_theta_r + q_cur[2]*diff_theta_b # q'k(\theta)
  inner_prod_v <- v_cur[1]*diff_theta_r + v_cur[2]*diff_theta_b # v'k(\theta)
  sign_inner_prod <- as.numeric(inner_prod_q > 0) # 1(q'k(\theta) > 0)
  
  score <- v_cur[1]*true_theta_01_rskew+v_cur[2]*true_theta_00_rskew+
    inner_prod_v*as.numeric(sign_inner_prod>0)
  
  # then the simulated support function in the current direction (q_cur, v_cur) is the sample average of score over X
  true_sf_rskew[i,] <- mean(score)
}


f <- function(c){ # objective function
  mean((-1-c)*true_theta_01_rskew+c*true_theta_00_rskew+
         ((-1-c)*diff_theta_r + c*diff_theta_b)*
         as.numeric((-1-c)*diff_theta_r + c*diff_theta_b>0))
}

c <- torch_tensor(0, requires_grad=TRUE)
optim <- optim_adam(params=list(c), lr=0.1)
for (i in 1:10000) {
  optim$zero_grad()            # initialize
  loss <- f(c)                 # get loss at current c
  loss$backward()              # backpropagation
  optim$step()                 # descent
}
true_F_coord_rskew <- -loss$item() # 0.354

plot_balance <- data.frame(e_r = true_sf_balance[1:num_q],
                           e_b = true_sf_balance[(num_q+1):(2*num_q)])

plot_rskew <- data.frame(e_r = true_sf_rskew[1:num_q],
                         e_b = true_sf_rskew[(num_q+1):(2*num_q)])

plot_RB_balance <- data.frame(e_r = c(true_sf_balance[(2*num_q)+1], true_sf_balance[(2*num_q)+3], true_F_coord_balance, logit_e_balance[1]), 
                              e_b = c(true_sf_balance[(2*num_q)+2], true_sf_balance[(2*num_q)+4], true_F_coord_balance, logit_e_balance[2]), 
                              type = c("true R", "true B", "true F", "alg_logit"))

plot_RB_rskew <- data.frame(e_r = c(true_sf_rskew[(2*num_q)+1], true_sf_rskew[(2*num_q)+3], true_F_coord_rskew, logit_e_rskew[1]),
                            e_b = c(true_sf_rskew[(2*num_q)+2], true_sf_rskew[(2*num_q)+4], true_F_coord_rskew, logit_e_rskew[2]),
                            type = c("true R", "true B", "true F", "alg_logit"))

```

```{r echo=FALSE, fig.width=8, fig.height=3, warning=FALSE}
write.csv(plot_RB_balance, "../truth_balance.csv", row.names = F)
write.csv(plot_RB_rskew, "../truth_rskew.csv", row.names = F)

print("Group-balanced config:")
print(paste0("true R=(", plot_RB_balance[1,1], ", ", plot_RB_balance[1,2], ")"))
print(paste0("true B=(", plot_RB_balance[2,1], ", ", plot_RB_balance[2,2], ")"))
print(paste0("true F=(", plot_RB_balance[3,1], ", ", plot_RB_balance[3,2], ")"))
print(paste0("`alg_logit` risk=(", plot_RB_balance[4,1], ", ", plot_RB_balance[4,2], ")"))

print("r-skewed config:")
print(paste0("true R=(", plot_RB_rskew[1,1], ", ", plot_RB_rskew[1,2], ")"))
print(paste0("true B=(", plot_RB_rskew[2,1], ", ", plot_RB_rskew[2,2], ")"))
print(paste0("true F=(", plot_RB_rskew[3,1], ", ", plot_RB_rskew[3,2], ")"))
print(paste0("`alg_logit` risk=(", plot_RB_rskew[4,1], ", ", plot_RB_rskew[4,2], ")"))

p1 <- ggplot(plot_balance, aes(x=e_r, y=e_b)) +
  geom_polygon(aes(x=e_r, y=e_b), colour='pink', fill='lavenderblush', alpha = 0.5) + 
  xlim(c(.15, .85)) + ylim(c(.2, .75)) +
  theme_classic() +
  geom_point(data = plot_RB_balance,
             aes(x = e_r, y = e_b, color=type, shape=type), 
             size=2.5) +
  scale_shape_manual("",values=c("alg_logit"=8, "true F"=13,
                                 "true R"=13, "true B"=13)) +
  scale_color_manual("", values=c("alg_logit"="darkgreen",
                                  "true R"="red", 
                                  "true B"="deepskyblue", 
                                  "true F"="black")) +
  geom_abline(intercept=0, slope=1, linetype="dashed", color="lightgray", linewidth=.25) +
  labs(title = "Group-balanced feasible set",
       x=TeX("$e_r$"),
       y=TeX("$e_b$"))

p2 <- ggplot(plot_rskew, aes(x=e_r, y=e_b)) +
  geom_polygon(aes(x=e_r, y=e_b), colour='pink', fill='lavenderblush', alpha = 0.5) +
  xlim(c(.15, .85)) + ylim(c(.2, .75)) +
  theme_classic() +
  geom_point(data = plot_RB_rskew,
             aes(x = e_r, y = e_b, color=type, shape=type),
             size=2.5) +
  scale_shape_manual("", values=c("alg_logit"=8, "true F"=13,
                                  "true R"=13, "true B"=13)) +
  scale_color_manual("", values=c("alg_logit"="darkgreen",
                                  "true R"="red", 
                                  "true B"="deepskyblue", 
                                  "true F"="black")) +
  geom_abline(intercept=0, slope=1, linetype="dashed", color="lightgray", linewidth=.25) +
  labs(title="r-skewed feasible set",
       x=TeX("$e_r$"),
       y=TeX("$e_b$"))
grid.arrange(p1, p2, ncol=2)
```

## Building confidence set for the fairness-accuracy frontier

The following code builds confidence sets for the fairness-accuracy frontier.

```{r}
n <- 10000
sample_frontier <- testDGP(n)

## group-balanced DGP -------
# create a grid of candidate values of e to test for
est_nuisance_balance <- nuisance(Y=sample_frontier$Y_balance,
                                 X=sample_frontier$X,
                                 G=sample_frontier$G)
cand_e_balance <- create_candidate_e_frontier(Y=sample_frontier$Y_balance,
                                              X=sample_frontier$X,
                                              G=sample_frontier$G,
                                              est_nuisance=est_nuisance_balance,
                                              num_grid=5000,
                                              buffer=2*sqrt(log(n))/sqrt(n))

# compute the test statistics, critical values, and rejection decision
FAfrontier_gs_balance <- CS_FAfrontier(Y=sample_frontier$Y_balance,
                                       X=sample_frontier$X,
                                       G=sample_frontier$G,
                                       e1=cand_e_balance[,1],
                                       e2=cand_e_balance[,2],
                                       est_nuisance=est_nuisance_balance,
                                       optimizer="grid", kink=TRUE,
                                       num_bstp_rep=1000)

plot_FAfrontier_balance <- concaveman(cand_e_balance[FAfrontier_gs_balance$rej==0,], concavity=4)

## r-skewed DGP -------
# create a grid of candidate values of e to test for
est_nuisance_rskew <- nuisance(Y=sample_frontier$Y_rskew,
                               X=sample_frontier$X,
                               G=sample_frontier$G)
cand_e_rskew <- create_candidate_e_frontier(Y=sample_frontier$Y_rskew,
                                            X=sample_frontier$X,
                                            G=sample_frontier$G,
                                            est_nuisance=est_nuisance_rskew,
                                            num_grid=5000,
                                            buffer=2*sqrt(log(n))/sqrt(n))

# compute the test statistics, critical values, and rejection decision
FAfrontier_gs_rskew <- CS_FAfrontier(Y=sample_frontier$Y_rskew,
                                     X=sample_frontier$X,
                                     G=sample_frontier$G,
                                     e1=cand_e_rskew[,1],
                                     e2=cand_e_rskew[,2],
                                     est_nuisance=est_nuisance_rskew,
                                     optimizer="grid", kink=TRUE,
                                     num_bstp_rep=1000)

plot_FAfrontier_rskew <- concaveman(cand_e_rskew[FAfrontier_gs_rskew$rej==0,], concavity=4)
```

```{r echo=FALSE, fig.width=8, fig.height=3.5, warning=FALSE}
p3 <- ggplot(plot_balance, aes(x=e_r, y=e_b)) +
  geom_polygon(data=data.frame(e_r=plot_FAfrontier_balance[,1], e_b=plot_FAfrontier_balance[,2]), aes(x = e_r, y = e_b), color="black", fill="lightsteelblue3", linewidth=0.2, alpha=0.5) +
  geom_polygon(aes(x=e_r, y=e_b), colour='pink', fill='lavenderblush', alpha = 0.5) + 
  xlim(c(.15, .85)) + ylim(c(.2, .75)) +
  theme_classic() +
  geom_point(data = plot_RB_balance,
             aes(x = e_r, y = e_b, color=type, shape=type), 
             size=2.5) +
  scale_shape_manual("",values=c("alg_logit"=8, "true F"=13,
                                 "true R"=13, "true B"=13)) +
  scale_color_manual("", values=c("alg_logit"="darkgreen",
                                  "true R"="red", 
                                  "true B"="deepskyblue", 
                                  "true F"="black")) +
  geom_abline(intercept=0, slope=1, linetype="dashed", color="lightgray", linewidth=.25) +
  labs(title = "Group-balanced feasible set",
       x=TeX("$e_r$"),
       y=TeX("$e_b$"))

# store the legend
legend <- ggplotGrob(p3)$grobs[[which(sapply(ggplotGrob(p3)$grobs, function(x) x$name) == "guide-box")]]

p4 <- ggplot(plot_rskew, aes(x=e_r, y=e_b)) +
  geom_polygon(data=data.frame(e_r=plot_FAfrontier_rskew[,1], e_b=plot_FAfrontier_rskew[,2]), aes(x = e_r, y = e_b), color="black", fill="lightsteelblue3", linewidth=0.2, alpha=0.5) +
  geom_polygon(aes(x=e_r, y=e_b), colour='pink', fill='lavenderblush', alpha = 0.5) +
  xlim(c(.15, .85)) + ylim(c(.2, .75)) +
  theme_classic() +
  geom_point(data = plot_RB_rskew,
             aes(x = e_r, y = e_b, color=type, shape=type),
             size=2.5) +
  scale_shape_manual("", values=c("alg_logit"=8, "true F"=13,
                                  "true R"=13, "true B"=13)) +
  scale_color_manual("", values=c("alg_logit"="darkgreen",
                                  "true R"="red", 
                                  "true B"="deepskyblue", 
                                  "true F"="black")) +
  geom_abline(intercept=0, slope=1, linetype="dashed", color="lightgray", linewidth=.25) +
  labs(title="r-skewed feasible set",
       x=TeX("$e_r$"),
       y=TeX("$e_b$"))

grid.arrange(
  arrangeGrob(p3 + theme(legend.position = "none"), p4 + theme(legend.position = "none"), ncol=2),
  legend,
  ncol = 2,
  widths = c(3, 0.5)
)
```
